---
title: "Homework 3"
author: "Aakriti Gupta"
date: 2018-10-14
output: github_document 
---


```{r, include=FALSE}
library(tidyverse)
library(ggridges)
library(patchwork)
```

## **Problem 1**
###Importing the CSV and data cleaning

* Importing the BRFSS dataset from p8105.datasets.
* Cleaning variable names
* Focusing on the topic "Overall Health"
* Formatted the data to keep only the useful variables
* Arranged reponses ordered from Excellent to Poor

```{r, data_import, cache=TRUE}
library(p8105.datasets)
data(brfss_smart2010)
```

```{r}
brfss = brfss_smart2010 %>% 
  janitor::clean_names() %>% 
  filter(topic == "Overall Health") %>% 
  select(-class, -topic, -question, -sample_size, -(confidence_limit_low:geo_location)) %>% 
    mutate(response = fct_relevel(response, c("Excellent", "Very good", "Good", "Fair", "Poor"))) %>%
  arrange(response)
```

### Answering questions using the cleaned dataset

* Question 1: In 2002, which states were observed at 7 locations?

```{r}
brfss %>% 
  filter(year == 2002) %>% 
  group_by(locationabbr) %>% 
  summarise(location_count = n_distinct(locationdesc)) %>%
  filter(location_count == 7) %>% 
  knitr::kable(digits = 1)
```

In the year 2002, **3 states namely Connecticut, Florida and North Carolina**, were observed at 7 locations.

* Question 2: Make a “spaghetti plot” that shows the number of locations in each state from 2002 to 2010.

```{r}
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = 1.0,
  out.width = "90%"
)
theme_set(theme_bw() + theme(legend.position = "right"))

brfss %>%
  group_by(year, locationabbr) %>% 
  summarise(location_count = n_distinct(locationdesc)) %>%
  rename(states = locationabbr) %>% 
  ggplot(aes(x = year, y = location_count)) + geom_smooth(aes(color = states), se = FALSE) + 
  labs(
    title = "Number of locations in each state from 2002 to 2010",
    x = "Year",
    y = "Number of distinct locations",
    caption = "Data from the BRFSS package"
  )
```

