---
title: "Homework 3"
author: "Aakriti Gupta"
date: 2018-10-14
output: github_document 
---


```{r, include=FALSE}
library(tidyverse)
library(ggridges)
library(patchwork)
```

## **Problem 1**
###Reading and cleaning the dataset

* Importing the BRFSS dataset from p8105.datasets.
* Cleaning variable names
* Focusing on the topic "Overall Health"
* Formatted the data to keep only the useful variables
* Arranged reponses ordered from Excellent to Poor

```{r, data_import, cache=TRUE}
library(p8105.datasets)
data(brfss_smart2010)
```

```{r}
brfss = brfss_smart2010 %>% 
  janitor::clean_names() %>% 
  filter(topic == "Overall Health") %>% 
  select(-class, -topic, -question, -sample_size, -(confidence_limit_low:geo_location)) %>% 
    mutate(response = fct_relevel(response, c("Excellent", "Very good", "Good", "Fair", "Poor"))) %>%
  arrange(response)
```

### Answering questions using the cleaned dataset

* Question 1: In 2002, which states were observed at 7 locations?

```{r}
brfss %>% 
  filter(year == 2002) %>% 
  group_by(locationabbr) %>% 
  summarise(location_count = n_distinct(locationdesc)) %>%
  filter(location_count == 7) %>% 
  knitr::kable(digits = 1)
```

In the year 2002, **3 states namely Connecticut, Florida and North Carolina**, were observed at 7 locations.

* Question 2: Make a “spaghetti plot” that shows the number of locations in each state from 2002 to 2010.

```{r}
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = 1.0,
  out.width = "90%"
)
theme_set(theme_bw() + theme(legend.position = "right"))

brfss %>%
  group_by(year, locationabbr) %>% 
  summarise(location_count = n_distinct(locationdesc)) %>%
  rename(states = locationabbr) %>% 
  ggplot(aes(x = year, y = location_count)) + geom_smooth(aes(color = states), se = FALSE) + 
  labs(
    title = "Number of locations in each state from 2002 to 2010",
    x = "Year",
    y = "Number of distinct locations",
    caption = "Data from the BRFSS package"
  )
```

* Question 3: Make a table showing, for the years 2002, 2006, and 2010, the mean and standard deviation of the proportion of “Excellent” responses across locations in NY State.



* Question 4: For each year and state, compute the average proportion in each response category (taking the average across locations in a state). Make a five-panel plot that shows, for each response category separately, the distribution of these state-level averages over time.

```{r}
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_bw() + theme(legend.position = "bottom"))

brfss %>% 
  group_by(year, locationabbr, response) %>% 
  summarise(avg_response = mean(data_value)) %>% 
  rename(states = locationabbr) %>% 
  ggplot(aes(x = year, y = avg_response)) + geom_smooth(aes(color = states), se = FALSE) + 
  facet_grid(. ~ response) +
  labs(
    title = "State-level averages of response categories over time",
    x = "Year",
    y = "Average responses",
    caption = "Data from the BRFSS package"
  )
```

## **Problem 2**
###Reading and cleaning the dataset

```{r, cache=TRUE}
library(p8105.datasets)
data(instacart)

insta = instacart %>% 
  janitor::clean_names()
```

###Summarizing the dataset
We will examine the instacart dataset. Instacart is an online grocery service that allows you to shop online from local stores.This dataset has `r nrow(insta)` observations and `r ncol(insta)` variables. The variables `aisle`, `department`, `eval_set` and `product_name` are of class `character`, whereas  all remaining variables are of class `integer`. 

###Answering questions
* Question 1: How many aisles are there, and which aisles are the most items ordered from?

**There are `r insta %>% pull(aisle) %>% n_distinct()` aisles in this dataset**.

Most items are ordered from **aisle 83 for fresh vegetables, aisle 24 for fresh fruits and aisle 123 for packaged vegetables fruits** as seen below.

```{r}
insta %>%
  group_by(aisle_id, aisle) %>% 
  summarise(item_count = n()) %>% 
  arrange(item_count) %>% 
  tail(3) %>% 
  knitr::kable()
```

* Question 2: Make a plot that shows the number of items ordered in each aisle. Order aisles sensibly, and organize your plot so others can read it.

```{r}
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = 0.6,
  out.width = "90%"
)
theme_set(theme_bw() + theme(legend.position = "bottom"))

insta %>%
  group_by(aisle_id, aisle) %>% 
  summarise(item_count = n()) %>%
  ggplot(aes(x = aisle_id, y = item_count)) + geom_point(color = "blue") + 
  labs(
    title = "Plot of number of items ordered in each aisle",
    x = "Aisle ID",
    y = "Number of items ordered",
    caption = "Data from the instacart package"
  ) +
   geom_text(aes(label = ifelse(item_count > 35000, as.character(aisle),'')), hjust = 1.1, vjust = 1.1)
```

* Question 3: Make a table showing the most popular item in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”.

```{r}
insta %>% 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>% 
  group_by(aisle, product_name) %>% 
  summarise(product_count = n()) %>%
  filter(product_count == max(product_count)) %>% 
  select(-product_count) %>% 
  knitr::kable()

```

